NGO analysis

### 2022-08-01 Dashboard/report.Rmd
Updating with two Excel functions; adding zero tests and warnings.

## Function to read Excel file -------------------------------------------------
widen_excel <- function(file) {
  read_excel(file) |>
    # uniting Category and Name to deal with multiple 'Other'
    unite(
      entry, c(Category, Name),
      sep = "_"
    ) |>
    # deleting Sub-category and Firm
    select(-`Sub-category`, -Firm) |>
    # pivoting to create 'tidy' variables
    pivot_wider(names_from = entry, values_from = Value) |>
    # cleaning names
    clean_names()
}
### Reading html ---------------------
# Creating the main data frame
htmldata <- read_proff(
  params$htmlfile,
  as.character(params$start), as.character(params$end)
)

@@

### Reading excel ---------------------
exceldata <- widen_excel(params$excelfile)

@@

NGO <- read_excel(params$excelfile) |>
  # Turning the needed column into a list
  pull(Firm) |>
  # discarding all NAs in case the name is somewhere
  purrr::discard(is.na) |>
  # choosing the first available name
  first()

@@

```{r data_warnings}
# Excel data tests =============================================================
# Functions to test for Excel data quality (basic)
datatest <-
  # let me know if anyone has a better idea for testing
  if (mean(unlist(lapply(exceldata, is.character))) != 0) {
    TRUE
  } else {
    FALSE
  }

## Creating the Excel warning text --------------------------------------------
excel_warning <- if (datatest == TRUE) {
  "***    Warning: something is wrong with the Excel data.***"
} else {
  ""
}

# HTML warnings ===============================================================
# Creating the arsresultat warning text
arsresultat_plot_warning <- if (sum(!is.na(htmldata$arsresultat)) == 0) {
  "*    No information about net income/loss is available.*"
} else {
  ""
}

@@

```{r functions_reliant}
## scale maximum--------------------------------------------------------------
# Function to create the maximum value for revenue and expenses plots
maximum <- if (datatest == TRUE) {
  NA
} else {
  # reading the Excel sheet anew
  read_excel(params$excelfile) |>
    # removing the firm and sub-category columns
    select(-Firm, -`Sub-category`) |>
    # filtering to retain only revenue and expenses entries
    filter(Category == "Revenue" | Category == "Expenses") |>
    # grouping and summarising to calculate maximum value
    group_by(Year, Category) |>
    summarise(
      sum = sum(Value, na.rm = TRUE)
    ) |>
    # Creating list of sums
    pull() |>
    # selecting the highest value
    max()
}

@@

```{r KPIs}
## Fundraising ROI------------------------------------------------------------
# Creating Fundraising ROI value
# testing for data quality, returning NA if bad
fundraising <- if (datatest == TRUE) {
  NA
} else {
  {
    exceldata |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (sum(revenue_private_contributions,
            revenue_membership_fees,
            na.rm = TRUE
          ) - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  } |> KPI_to_string()
}

@@

@@

# Printing the plot if the data are OK
if (datatest == FALSE) {
  revenue_plot
}

@@

# Printing the plot if the data are OK
if (datatest == FALSE) {
  expenses_plot
}
```

>>>>

### Reading html ---------------------
# Creating the main data frame
htmldata <- read_proff(
  params$htmlfile,
  as.character(params$start), as.character(params$end)
)

@@

## Functions to read Excel file ------------------------------------------------
### First Excel transformation -------------------------------------------------
# creating df used by NGO title function and maximum scale function
widen_excel_test <- function(file) {
  # Reading Excel file
  read_excel(file) |>
    # cleaning names to enable later operations
    clean_names()
}

### Second Excel function - creating primary Excel data frame-------------------
widen_excel <- function(object) {
  object |>
    # uniting category and name to deal with multiple 'Other'
    unite(
      entry, c(category, name),
      sep = "_"
    ) |>
    # deleting sub_category and firm
    select(-sub_category, -firm) |>
    # pivoting to create 'tidy' variables
    pivot_wider(names_from = entry, values_from = value) |>
    # cleaning names
    clean_names()
}

### Reading excel ---------------------
excel_test <- widen_excel_test(params$excelfile)

exceldata <- widen_excel(excel_test)

@@

NGO <- excel_test |>
  # Turning the needed column into a list
  pull(firm) |>
  # discarding all NAs in case the name is somewhere
  purrr::discard(is.na) |>
  # choosing the first available name
  first()

@@

```{r data_warnings}
# Excel data tests and warnings ================================================
# Functions to test for Excel data quality (basic)
## Excel datatest---------------------------------------------------------------
datatest <-
  # let me know if anyone has a better idea for testing
  if (mean(unlist(lapply(exceldata, is.character))) != 0) {
    TRUE
  } else {
    FALSE
  }

## Excel zero test------------------------------------------------------------
excel_zero_check <- function(type) {
  excel_test |>
    filter(
      category == as.character(all_of(type))
    ) |>
    summarise(
      zero = sum(value != 0, na.rm = TRUE)
    ) |>
    pull()
}

## Creating the Excel warning texts --------------------------------------------
# Excel data test warning
excel_warning <- if (datatest == TRUE) {
  "***    Warning: something is wrong with the Excel data.***"
} else {
  ""
}

# Excel zero revenues warning
excel_zero_revenue_warning <- if (excel_zero_check("Revenue") == 0) {
  "*    All revenue values are zero or missing. This is unusual and may indicate a data problem.*"
} else {
  ""
}

# Excel zero expenses warning
excel_zero_expenses_warning <- if (excel_zero_check("Expenses") == 0) {
  "*    All expenses values are zero or missing. This is unusual and may indicate a data problem.*"
} else {
  ""
}

# HTML warnings ================================================================
# Creating the arsresultat warning text
arsresultat_plot_warning <- if (sum(!is.na(htmldata$arsresultat)) == 0) {
  "*    No information about net income/loss is available.*"
} else {
  ""
}

@@

```{r functions_reliant}
## scale maximum--------------------------------------------------------------
# Function to create the maximum value for revenue and expenses plots
maximum <- if (datatest == TRUE) {
  NA
} else {
  excel_test |>
    # removing the firm and sub-category columns
    select(-firm, -sub_category) |>
    # filtering to retain only revenue and expenses entries
    filter(category == "Revenue" | category == "Expenses") |>
    # grouping and summarising to calculate maximum value
    group_by(year, category) |>
    summarise(
      sum = sum(value, na.rm = TRUE)
    ) |>
    # Creating list of sums
    pull() |>
    # selecting the highest value
    max()
}

@@

```{r KPIs}
## Fundraising ROI------------------------------------------------------------
# Creating Fundraising ROI value
# testing for data quality, returning NA if bad
fundraising <- if (datatest == TRUE) {
  NA
} else {
  exceldata |>
    # most recent year
    filter(year == max(year)) |>
    transmute(
      # Calculating the value as a percentage; removing all other
      KPI_fundraising_num =
        (sum(revenue_private_contributions,
          revenue_membership_fees,
          na.rm = TRUE
        ) - expenses_fundraising) /
          expenses_fundraising * 100
    ) |>
    # just want the single number
    pull() |>
    first() |>
    KPI_to_string()
}

@@

`r excel_zero_revenue_warning`

`r excel_zero_expenses_warning`

@@

# Printing the plot if the data are OK
if (datatest == FALSE && excel_zero_check("Revenue") != 0) {
  revenue_plot
}

@@

# Printing the plot if the data are OK
if (datatest == FALSE && excel_zero_check("Expenses") != 0) {
  expenses_plot
}
```
